Marker_identification_Benchmark

### 2024-03-26 Demo_marrow_facs.Rmd
24/03/26

```{r}
tic()
lmd_result = LMD(dat, feature_space)
toc()
```

@@

### Annotate Cell Cycle
```{r}
mmus_s = gprofiler2::gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
mmus_g2m = gprofiler2::gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
tiss = CellCycleScoring(tiss, s.features = mmus_s, 
                        g2m.features = mmus_g2m, set.ident = FALSE)
cell_cycle_table = data.frame(tiss@meta.data[,c("S.Score","G2M.Score","Phase")],"Dataset" = "FACS")
```

>>>>

```{r}
tic()
lmd_result = LMD(dat, feature_space)
toc()
```
```{r}
lmd_result = LMD(dat, feature_space, highres = TRUE, max_time = 2^7)
```

@@

### Annotate Cell Cycle
```{r}
mmus_s = gprofiler2::gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
mmus_g2m = gprofiler2::gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
tiss = CellCycleScoring(tiss, s.features = mmus_s, 
                        g2m.features = mmus_g2m, set.ident = FALSE)
tiss <- tiss %>%
  ScaleData(verbose = FALSE, features = c(mmus_s,mmus_g2m)) %>% RunPCA(npcs = 20, verbose = FALSE, features = c(mmus_s,mmus_g2m)) %>% RunTSNE(dims = 1:20, seed.use = 42, reduction.name = "tsne_cc")

cell_cycle_table = data.frame(tiss@meta.data[,c("S.Score","G2M.Score","Phase")],"Dataset" = "FACS")
```

### 2024-03-26 Marker_identfication_benchmark.Rmd
24/03/26

library("Seurat")
library("dplyr")
```

# Load Data & Process
## Load Data
```{r}
# tiss <- readRDS("/data/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
```

@@

if(tissue_fullname == "Bone marrow"){
  tissue_name = "marrow_facs"
  if (!file.exists('facs_Marrow_seurat_tiss.Robj')) {
   # If load url doesn't work, please download it and load from your working directory
  options(timeout=6000)
  download.file("https://figshare.com/ndownloader/files/13092380",
                destfile = 'facs_Marrow_seurat_tiss.Robj', method = 'libcurl')
  }
  load('facs_Marrow_seurat_tiss.Robj')
  tiss <- UpdateSeuratObject(tiss)
}if(tissue_fullname == "Pancreas"){
  tissue_name = "pancreas_facs"
  if (!file.exists('facs_Pancreas_seurat_tiss.Robj')) {
     # If load url doesn't work, please download it and load from your working directory
    options(timeout=6000)
    download.file("https://figshare.com/ndownloader/files/13092386",
                  destfile = 'facs_Pancreas_seurat_tiss.Robj', method = 'libcurl')
  }
  load('facs_Pancreas_seurat_tiss.Robj')
  tiss <- UpdateSeuratObject(tiss)
}if(tissue_fullname == "Lung"){
  tissue_name = "lung_facs"
  if (!file.exists('facs_Lung_seurat_tiss.Robj')) {
     # If load url doesn't work, please download it and load from your working directory
    options(timeout=6000)
    download.file("https://figshare.com/ndownloader/files/13092194",
                  destfile = 'facs_Lung_seurat_tiss.Robj', method = 'libcurl')
   }
  load('facs_Lung_seurat_tiss.Robj')
  tiss <- UpdateSeuratObject(tiss)
}if(tissue_fullname == "Thymic"){
  options(timeout=6000)
  if (!file.exists('Human_Thymic_Epithelial_seurat_tiss.Robj')) {
     # If load url doesn't work, please download it and load from your working directory
    options(timeout=6000)
    download.file("https://datasets.cellxgene.cziscience.com/046d2e35-2ee6-4159-8db7-7b1039fe0005.rds",
                  destfile = 'Human_Thymic_Epithelial_seurat_tiss.Robj', method = 'libcurl')
  }
  tiss <- readRDS()
}

tissue_fullname = "Bone marrow"
```
## Process
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
cell_label = tiss$cell_ontology_class

@@

# Define Ground truth Marker
## Criterion1: Fold-change
For each gene, sorting cell types by mean expression for each gene and computing fold changes between consecutive types, take the maximum value among the N-1 fold change values, given N cell types.
```{r}
folder_path = "benchmark_result"
# Check if the folder exists
if (!dir.exists(folder_path)) {
  # If it doesn't exist, create it
  dir.create(folder_path)
} else {
}
```

@@

# Run Each Method and save results
## LocalMarkerDetector
Genes were ranked by the increasing order of Cumulative Diffusion-KL score
```{r}
method = "lmd"
source("LMD_function.R", echo = F)
res = LMD(dat,feature_space,max_time = 2^10,knn = 5)
marker = show_result_lmd(res, n = length(selected_genes))
marker[order(marker$score),'rank'] = 1:nrow(marker)
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```

## Seurat v4 Wilcoxon unfiltered
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)
method = "wilcox_no_filter"
marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
gene_rank$'rank' = 1:nrow(gene_rank)
marker <- merge(marker,gene_rank,by = "gene")
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```
## Seurat v4 default
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)

@@

## Marcopolo
The rank of genes is given by `Marcopolo_rank`
```{python}
import pickle
import numpy as np
import pandas as pd
import torch
import anndata as ad
import scanpy as sc
import matplotlib.pyplot as plt
import MarcoPolo

@@

result = MarcoPolo.find_markers(adata=adata, regression_result=regression_result)
```
```{r}
method = "marcopolo"
marker = reticulate::py$result
marker$'rank' = marker$MarcoPolo_rank + 1
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```

@@

# Integrate rank from all methods into one dataframe
```{r}
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")

df_benchmark = do.call(cbind,lapply(method_ls, function(method){
  marker = read.table(file.path(folder_path, paste0(method,"_",tissue_name,".csv")),header = TRUE)
  if("gene" %in% colnames(marker)){
    return(marker$rank[match(selected_genes,marker$gene)])
  }else{
    return(marker$rank[match(selected_genes,rownames(marker))])
  }
}))
df_benchmark[is.na(df_benchmark)] = nrow(df_benchmark)
df_benchmark = data.frame(df_benchmark,row.names = selected_genes)
colnames(df_benchmark) = method_ls

write.table(df_benchmark,file = file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")),row.names = TRUE)
```

@@

method_ls = c("lmd","lmd_gaussian","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
```
```{r}
max_logfc = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c1.txt"))) %>% rownames()
celldb_marker = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c2.txt")))[,1]
gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
names(gt_list) = c(paste0("Top",seq(50,1000,50)),"CellMarkerDB")

@@

# Density Index
```{r}
df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))

score_measure =do.call(rbind,lapply(c(method_ls,"All genes"),function(method){
  if(method == "All genes"){
    top_gs = rownames(df_benchmark)
    top_n = length(top_gs)
  }else{
    top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
    top_n = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,max(df_benchmark),2000))
  }
  result = unlist(lapply(top_n, function(top){
    tiss <- RunPCA(tiss, features = top_gs[1:top], npcs = 20)
    d = dist(Embeddings(tiss, reduction = "pca"))
    c = as.integer(factor(tiss$cell_ontology_class[labels(d)]))

    # # Silhouette score
    # s1 = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
    #   group_by(cluster) %>%
    #   summarize(mean_silhouette = mean(sil_width)) %>%.$mean_silhouette)

    # Density-index
    nn.dist = FNN::get.knn(Embeddings(tiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
    sdVec <- stats::na.omit(tiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2))
    s2 = length_scale/mean(nn.dist)

@@

  }))
  # rownames(result) = c("Mean_SI","DI")
  result = data.frame(TopGene = top_n,Method = method,DensityIndex = result)
}) )

write.table(score_measure,file = file.path(folder_path, paste0(tissue_name,"_density_index.csv")),row.names = FALSE)
```

>>>>

library("Seurat")
library("dplyr")
library(RColorBrewer)
```

# Load Data & Process
## Load Data
```{r}
# tiss <- readRDS("/data/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
# DefaultAssay(tiss) <- "RNA"
```

@@

if(tissue_fullname == "Bone marrow"){
  tissue_name = "marrow_facs"
  if (!file.exists('facs_Marrow_seurat_tiss.Robj')) {
   # If load url doesn't work, please download it and load from your working directory
  options(timeout=6000)
  download.file("https://figshare.com/ndownloader/files/13092380",
                destfile = 'facs_Marrow_seurat_tiss.Robj', method = 'libcurl')
  }
  load('facs_Marrow_seurat_tiss.Robj')
  tiss <- UpdateSeuratObject(tiss)
}
if(tissue_fullname == "Pancreas"){
  tissue_name = "pancreas_facs"
  if (!file.exists('facs_Pancreas_seurat_tiss.Robj')) {
     # If load url doesn't work, please download it and load from your working directory
    options(timeout=6000)
    download.file("https://figshare.com/ndownloader/files/13092386",
                  destfile = 'facs_Pancreas_seurat_tiss.Robj', method = 'libcurl')
  }
  load('facs_Pancreas_seurat_tiss.Robj')
  tiss <- UpdateSeuratObject(tiss)
}
if(tissue_fullname == "Lung"){
  tissue_name = "lung_facs"
  if (!file.exists('facs_Lung_seurat_tiss.Robj')) {
     # If load url doesn't work, please download it and load from your working directory
    options(timeout=6000)
    download.file("https://figshare.com/ndownloader/files/13092194",
                  destfile = 'facs_Lung_seurat_tiss.Robj', method = 'libcurl')
   }
  load('facs_Lung_seurat_tiss.Robj')
  tiss <- UpdateSeuratObject(tiss)
}
if(tissue_fullname == "Thymic"){
  options(timeout=6000)
  if (!file.exists('Human_Thymic_Epithelial_seurat_tiss.Robj')) {
     # If load url doesn't work, please download it and load from your working directory
    options(timeout=6000)
    download.file("https://datasets.cellxgene.cziscience.com/046d2e35-2ee6-4159-8db7-7b1039fe0005.rds",
                  destfile = 'Human_Thymic_Epithelial_seurat_tiss.Robj', method = 'libcurl')
  }
  load('Human_Thymic_Epithelial_seurat_tiss.Robj')
}
```

## Process
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = pmin(dim(tiss@reductions$pca@cell.embeddings)[2],30)
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)

@@

# Define Ground truth Marker
## Criterion1: Fold-change
For each gene, sorting cell types by mean expression for each gene and computing fold changes between consecutive types, take the maximum value among the N-1 fold change values, given N cell types.
```{r}
folder_path = "benchmark_result"
# Check if the folder exists
if (!dir.exists(folder_path)) {
  # If it doesn't exist, create it
  dir.create(folder_path)
}
```

@@

# Run Each Method and save results
## LocalMarkerDetector
Genes were ranked by the increasing order of Cumulative Diffusion-KL score
```{r}
method = "lmd"
source("LMD_function.R", echo = F)
res = LMD(dat,feature_space,max_time = 2^20,knn = 5)
marker = show_result_lmd(res, n = length(selected_genes))
marker[order(marker$score),'rank'] = 1:nrow(marker)
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```

## Seurat v4 Wilcoxon unfiltered
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)
method = "wilcox_no_filter"
# marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
# gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
marker <- presto:::wilcoxauc.Seurat(X = subset(tiss,features = selected_genes),
                                    assay = "data", 
                                    seurat_assay = "RNA")
colnames(marker)[1] = "gene"
gene_rank = marker %>% arrange(.,padj,desc(logFC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
gene_rank$'rank' = 1:nrow(gene_rank)
marker <- merge(marker,gene_rank,by = "gene")
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```
## Seurat v4 default
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)

@@

## Marcopolo
Long Runtime Option
```{r}
# transfer the seurat obj to h5ad & process in jupyter notebook
library(reticulate)
library(anndata)
sc <- import("scanpy", convert = FALSE)
scvi <- import("scvi", convert = FALSE)
# scvi$settings$progress_bar_style = 'tqdm'
tiss_adata <- sc$AnnData(
  X = t(raw_count)
)
anndata::write_h5ad(tiss_adata, "./dataset/tm_marrow.h5ad")
```
The rank of genes is given by `Marcopolo_rank`
```{python}
import pickle
import numpy as np
import pandas as pd
import torch
import anndata as ad
import scanpy as sc
import matplotlib.pyplot as plt
import MarcoPolo

@@

result = MarcoPolo.find_markers(adata=adata, regression_result=regression_result)
result = result.set_axis(list(adata.var.index) ,axis = 0)
```
```{r}
method = "marcopolo"
marker = reticulate::py$result
marker$'rank' = marker$MarcoPolo_rank + 1
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```

@@

# Integrate rank from all methods into one dataframe
```{r}
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
df_benchmark = lapply(method_ls, function(method){
  marker = read.table(file.path(folder_path, paste0(method,"_",tissue_name,".csv")),header = TRUE)
  if(!"gene" %in% colnames(marker)){
    marker$'gene' = rownames(marker)
  }
  marker %>% select(gene,rank) %>% distinct()
})
names(df_benchmark) = method_ls
df_benchmark = bind_rows(df_benchmark, .id = "method")
df_benchmark <- tidyr::pivot_wider(df_benchmark, names_from = method, values_from = rank)
df_benchmark[is.na(df_benchmark)] = nrow(df_benchmark)

# rownames(df_benchmark) = df_benchmark$gene
# df_benchmark[order(Gene_detected_count[rownames(df_benchmark)]),"gene_count"] = 1:nrow(df_benchmark)

write.table(df_benchmark,file = file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")),row.names = FALSE)
```

@@

method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
```
```{r}
max_logfc = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c1.txt"))) %>% rownames()
celldb_marker = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c2.txt")))[,1]
gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
names(gt_list) = c(paste0("Top",seq(50,1000,50)),"CellMarkerDB")

@@

# Density Index
```{r}
folder_path = "benchmark_result"
df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))
score_measure =do.call(rbind,lapply(c(colnames(df_benchmark),"All genes"),function(method){
  if(method == "All genes"){
    top_gs = rownames(df_benchmark)
    top_n = length(top_gs)
  }else{
    top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
    # top_n = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,max(df_benchmark),2000))
    top_n = c(seq(50,90,10),seq(100,1000,100))
  }
  DI = unlist(lapply(top_n, function(top){
    features = top_gs[1:top]
    # remove drop-out cells
    subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
    subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)
    # # Silhouette score
    # d = dist(Embeddings(subtiss, reduction = "pca"))
    # c = as.integer(factor(subtiss$cell_ontology_class[labels(d)]))
    # s1 = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
    #   group_by(cluster) %>%
    #   summarize(mean_silhouette = mean(sil_width)) %>%.$mean_silhouette)

    # Density-index
    nn.dist = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
    s2 = length_scale/mean(nn.dist)

@@

  }))
  # rownames(result) = c("Mean_SI","DI")

  # Non-dropout Rate
  cell_count = lapply(top_n,function(n){
    colSums(dat[top_gs[1:n],]>0)
  })
  non_dropout_rate = unlist(lapply(cell_count,function(x){
    sum(x != 0)/length(x)
  }))

  result = data.frame(TopGene = top_n,Method = method,DensityIndex = DI, non_dropout_rate = non_dropout_rate)
  result
}) )

write.table(score_measure,file = file.path(folder_path, paste0(tissue_name,"_density_index_no_dropout.csv")),row.names = FALSE)

# error bar & null distribution
top_n = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,max(df_benchmark),2000))
null_score_measure = do.call(rbind,lapply(top_n,function(top){
  set.seed(top)
  seed_ls = sample(1:1e5,size = 100)
  random_top_gs = lapply(seed_ls,function(sed){
    set.seed(sed)
    sample(rownames(df_benchmark),top,replace = FALSE)
  })
  result = unlist(lapply(random_top_gs, function(g_ls){
    features = g_ls
    subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
    subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)
    # Density-index
    nn.dist = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2))
    s2 = length_scale/mean(nn.dist)
    s2
  }))
  result = data.frame(TopGene = top,DensityIndex = result)
}) )
write.table(null_score_measure,file = file.path(folder_path, paste0(tissue_name,"_DI_null.csv")),row.names = FALSE)

# smallest set of genes covering 95% cells & corresponding DI
min_coverage = 1
coverage_score_measure = do.call(rbind,lapply(colnames(df_benchmark),function(method){
  cat(method,"\n")
  top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
  min_covered = apply(dat[top_gs,] > 0, 2, function(x) which(x)[min_coverage])
  min_row_for_coverage <- as.integer(names(which(cumsum(table(min_covered)) >= ncol(dat) * 0.95)[1]))
  features = top_gs[1:min_row_for_coverage]

  # remove drop-out cells
  subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
  npcs = pmin(20,length(features))
  subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = npcs, verbose = FALSE)
  subtiss <- RunUMAP(subtiss,dims = 1:dim(subtiss@reductions$pca)[2])
  # Density-index
  if(length(features) < npcs){
    nn.dist = FNN::get.knn(t(subtiss[["RNA"]]@scale.data), k = 10, algorithm = "kd_tree")$nn.dist
    length_scale <- norm(t(subtiss[["RNA"]]@scale.data),type = "F")/sqrt(ncol(subtiss)-1)
    s2 = length_scale/mean(nn.dist)
  }else{
    nn.dist = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
    s2 = length_scale/mean(nn.dist)
  }

  data.frame(TopGene = min_row_for_coverage,Method = method,DensityIndex = s2, non_dropout_rate = sum(colSums(dat[features,]>0)!=0)/ncol(dat))

  DimPlot(subtiss,group.by = "cell_ontology_class") + NoLegend() + ggtitle(sprintf("%s, %d genes, DI: %.4f",method,min_row_for_coverage,s2))
}))

```

# NNret + Silhouette + DI
```{r}
folder_path = "benchmark_result"
df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))
score_measure =do.call(rbind,lapply(c(colnames(df_benchmark),"All genes"),function(method){
  if(method == "All genes"){
    top_gs = rownames(df_benchmark)
    top_n = length(top_gs)
  }else{
    top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
    top_n = c(seq(50,90,10),seq(100,1000,100))
  }
  scores = sapply(top_n, function(top){
    features = top_gs[1:top]
    # remove drop-out cells
    subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
    subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)

    # Silhouette score
    d = dist(Embeddings(subtiss, reduction = "pca"))
    c = as.integer(factor(subtiss$celltype[labels(d)]))
    si = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
      group_by(cluster) %>%
      summarize(mean_silhouette = mean(sil_width)) %>%.$mean_silhouette)

    # NNret
    nn.result = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")
    nn.dist = nn.result$nn.dist
    nn.index = nn.result$nn.index
    nnret = mean(
      unlist(lapply(1:nrow(nn.index),function(i){
      mean(c[nn.index[i,]] == c[i])
    })) )

    # Density-index
    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
    di = length_scale/mean(nn.dist)

    c(si,di,nnret)
  }) 
  scores = t(scores)
  colnames(scores) = c("Sihouette_score","DensityIndex","NNret")
  rownames(scores) = top_n

  # Non-dropout Rate
  cell_count = lapply(top_n,function(n){
    colSums(tiss[["RNA"]]@data[top_gs[1:n],]>0)
  })
  non_dropout_rate = unlist(lapply(cell_count,function(x){
    sum(x != 0)/length(x)
  }))

  result = data.frame(TopGene = top_n,Method = method, scores, non_dropout_rate = non_dropout_rate)
  result
}) )
```

# Recovery of celltypes
```{r}
# jaccard index of cell-type-specificity of each genes
cell_label = tiss$celltype
one_hot_matrix_ct <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))
binary_dat = dat > apply(dat,2,median)
intersect_mtx = binary_dat %*% one_hot_matrix_ct
union_mtx = sapply(colnames(one_hot_matrix_ct),function(ct){
  rowSums((binary_dat > 0)) + colSums(one_hot_matrix_ct)[ct]
}) - intersect_mtx
jaccard_sim = intersect_mtx / union_mtx
threshold = 0.8 / ((ncol(tiss)/table(cell_label))*0.2 + 0.8)[colnames(jaccard_sim)]
recovery_mtx = t(apply(jaccard_sim,1,function(x) x > threshold))

# Prop of recovered cell types
recovery_rate =data.frame(sapply(c(colnames(df_benchmark)),function(method){
  top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
  top_n = 1:300
  scores = sapply(top_n, function(top){
    features = top_gs[1:top]
    mean(colSums(recovery_mtx[features,,drop = FALSE]) > 0)
  }) 
}) )
recovery_rate$'TopGene' = 1:nrow(recovery_rate)
```
